% Security of mixnets with a cascade topology.

\documentclass[letter]{article}
\usepackage{color,soul}
\usepackage[dvipsnames]{xcolor}
\usepackage[utf8]{inputenc}
\usepackage[margin=4cm]{geometry}
\usepackage{graphics}
\usepackage[toc,page]{appendix}
\usepackage[font={small}]{caption}
\usepackage{mathdots}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{parskip}
\usepackage{framed}
\usepackage{multicol}

\hypersetup{
    colorlinks,%
    citecolor=black,%
    filecolor=black,%
    linkcolor=black,%
    urlcolor=black
}

\def\dashuline{\bgroup
  \ifdim\ULdepth=\maxdimen  % Set depth based on font, if not set already
	  \settodepth\ULdepth{(j}\advance\ULdepth.4pt\fi
  \markoverwith{\kern.15em
	\vtop{\kern\ULdepth \hrule width .3em}%
	\kern.15em}\ULon}

\newcounter{foot}
\setcounter{foot}{1}
\setlength\parindent{2em}


%%%%% Macros. %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\DeclareMathAlphabet\mathbfcal{OMS}{cmsy}{b}{n}
\newcommand{\note}[1]{}

% Sets
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Zmodp}{\mathbb{Z}_p}
\newcommand{\Zmodq}{\mathbb{Z}_q}
\newcommand{\N}{\mathbb{N}}
\newcommand{\bits}{\{0,1\}}
\newcommand*\Union{\bigcup}
\newcommand*\Intersect{\bigcap}
\newcommand*\union{\cup}
\newcommand*\intersect{\cap}
\newcommand*\by{\times}
%\newcommand{\getsr}{\xleftarrow{\text{\tiny{\$}}}}
\newcommand{\getsr}{\mathrel{\leftarrow\mkern-14mu\leftarrow}}

% Asymptotics
\newcommand{\negl}{\proc{negl}}
\newcommand{\poly}{\proc{poly}}

\newcommand{\heading}[1]{\noindent \textsc{#1}}

% Misc.
\newcommand{\brackets}[1]{\langle #1 \rangle}
\newcommand{\view}{\mathsf{view}}
\newcommand{\class}[1]{\mathbf{#1}}
\newcommand{\Perm}{\textsf{Perm}}

\newcommand{\pk}{K}
\newcommand{\sk}{k}
\newcommand{\varm}{M}
\newcommand{\varc}{C}
\newcommand{\varr}{R}
\newcommand{\varx}{X}
\newcommand{\vary}{Y}

\newcommand{\advD}{D}
\newcommand{\advA}{A}

\newcommand{\gen}{\mathcal{G}}
\newcommand{\enc}{\mathcal{E}}
\newcommand{\dec}{\mathcal{D}}

%%%%% Title. %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\date{\today}
\title{Shuffling in secret (and proving it too)}
  \author{Chris Patton}

\setcounter{tocdepth}{2}

\renewcommand{\cal}[1]{\mathcal{#1}}

\newcommand{\oneCol}[2]{
\begin{center}
        \framebox{
        \begin{tabular}{c@{\hspace*{.4em}}}
        \begin{minipage}[t]{#1\textwidth} #2 \end{minipage}
        \end{tabular}
        }
\end{center}
}

%%%%% Content. %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

  \maketitle

\section{Introduction}

A mixnet (sometimes called a ``mix-network'') is a service designed to
facilitate anonymous communications over the internet \cite{chaum81}. Each of a
set of users $\cal{Q} = \{P_1, \dots, P_n\}$ wish to send a message to a user in
$\cal{Q}$. Each encrypts their message $M$ and the identity of the intended
recipient $P_i$ under the public key of the mixnet server (heereafter referred
to as the ``mix'') and transmits the ciphertext.  The mix server waits until it
receives $m \le n$ ciphertexts from $m$ distinct users.  It then decrypts each
message and destination pair under its secret key and transmits each message to
its intended destination. Crucially, the mix transmits the messages in an
unpredictable order so as to hide the correspondence between senders and
recipients.  As long as the mix is trusted and the user's ciphertexts are
faithfully transmitted to the mix, this simple approach suffices to preserve
anonymity of its participants.

Normally we think of the mixnet being composed of a cascade of mix servers, each
decrypting, shuffling, then transmitting to the next mix in the cascade until
finally the last server outputs the users' messages. This is meant to distribute
trust so that as long as the messages pass through at least one honest mix,
anonymity is achieved. Still, the trust model here is rather strong. For
example, the first mix could send a fake message on behalf of one of the
senders, or the last mix could redirect a message to a different recipient.  We
consider the problem of publicly verifying that the outputs are indeed a
permutation of the inputs. Moreover, we require that the correspondence not be
revealed. Andrew Neff describes in \cite{neff01} a method for proving a shuffle
in zero-knowledge.

Neff's solution is based on the ElGamal public key encryption scheme, which is
known to be semantically secure assuming the discrete log problem (decisional
Diffie-Hellman, or DDH) is hard \cite{elgamal84}. Based on this assumption, Neff
proves his protocol achieves a property known as honest verifier zero-knowledge
(HVZK) \cite{goldreich}. Informally speaking, this means that as long as the verifier
faithfully executes the protocol, the mix can prove the outputs are a
permutation of the inputs without revealing the permutation.

We formulate and define security for the problem and present Neff's protocol.
For simplicity, we limit the discussion to a mixnet with a single mix server. In
this scenario, users encrypt their messages under the public key of the mix. The
mix waits until it receives a full batch of ciphertexts, decrypts them under its
private key, and shuffles the output. In section~\ref{section:prelim}, we specify
ElGamal encryption, describe the DDH assumption, and define the HVZK notion. In
section~\ref{section:neff-shuffle}, we present the protocol and sketch its
security. In section~\ref{sec-implementation} we discuss the details of the
implementation.

\section{Preliminaries} \label{section:prelim}

If $X$ is a finite set, let $x \getsr X$ denote uniformly sampling an
element $x$ from the set. Let $y \gets A(x)$ denote running algorithm $A$ on input $x$ and
assigning its output to $y$. Let $y \gets A(x; r)$ denote the execution of a
probabilistic algorithm with the sequence of coins $r \in \bits^\infty$. Let $y
\getsr A(x)$ denote choosing $r \getsr \bits^\infty$ and executing $y \gets
A(x; r)$. Let $[i..j]$ where $i \le j$ denote the set of integers from $i$ to
$j$ inclusively.
Let $[n] = [1..n]$.
Let $\Perm(n)$ denote the set of permutations on vectors of length $n$.
Let $y \getsr \brackets{P, V}(x)$ denote the probabilistic execution of
interactive Turing machines $P$ and $V$ on common input $x$. When the protocol
finishes, the output of $V$ is assigned to $y$.
$x$ is called the common input of $P$ and $V$.
Finally a function $\Delta(n)$ is
negligible if for all positive polynomials
$p(\cdot)$ and sufficiently large $n$, $\Delta(n) < \frac{1}{p(n)}$.


\subsection{ElGamal encryption}

Fix a prime number $p$ and let $\Zmodp$ denote the modular ring of order $p$.
In what follows, we assume arithmetic is performed in the modular ring $\Zmodp$.
Let $\Zmodp^*$ denote the corresponding multiplicative group.
Let $g \in \Zmodp$ such that $g^q = 1$ where $q$ is prime. Then $\brackets{g} =
\Zmodq^*$ is a subgroup of $\Zmodp^*$ and $q | (p-1)$. We call $(p, q, g)$ the
\textit{public parameters}. The \textit{secret key} is chosen uniformly from
the set of powers of $g$ mod $p$. This is denoted by $\sk \getsr [q-1]$. Let $\pk = g^\sk$ be
the \textit{public key}. Key generation is denoted $(K, k) \getsr \gen(p, q,
g)$.
Encryption and decryption operate on group elements.\footnote{
Public-key encryption schemes are typically defined on bit
strings. It is easy to messages as elements of $\Zmodp^*$, but the number of
messages that can be encoded depends on the modulus $p$.}
Encryption of $M \in \Zmodp^*$ proceeds as follows:
let $r \getsr [q-1]$,
$R = g^r$,
$S = K^r$, and
$C = M \cdot S$.
Output the tuple $(C, R)$.
This is denoted $(C, R) \getsr \enc_K(M)$.
Decryption of a ciphertext $(C, R) \in \Zmodp^* \by \Zmodp^*$ proceeds as follows:
let $S = R^\sk$ ($S$ is called the \emph{shared secret}) and
$M = C \cdot S^{-1}$.
Normally the decryption algorithm would output just $M$, but for the purpose of
verification, we demand that it also returns the shared secret. Thus,
decryption is denoted $(M, S) \gets \dec_k(C, R)$.

ElGamal is known to be \textit{semantically secure} under the
decisional Diffie-Hellman assumption \cite{elgamal84,ty98}. Informally speaking, this
means that if given $\pk$, it is infeasible for any adversary to determine
$\sk$, then given $(C,R)$, it is infeasible for any adversary to
learn anything meaningful about $M$. As we're not
interested in the security of ElGamal from a privacy standpoint, we won't
formalize semantic security. However, we define the DDH assumption in the next
subsection.

The mixnet operates as follows:
let $\varm_1, \ldots, \varm_n \in \Zmodp^*$ be a sequence of plaintexts.
Let $(\varc_1, \varr_1), \ldots, (\varc_n, \varr_n) \in \Zmodp^* \by \Zmodp^*$ be a
sequence of corresponding ElGamal ciphertexts encrypted under the public-key
$\pk$. The mix server, who possesses the secret-key $\sk$, takes as input
the ciphertexts and outputs a sequence $(\varm_{\pi(1)}, S_{\pi(1)}), \ldots,
(\varm_{\pi(n)}, S_{\pi(n)})$.
where $\pi$ is a permutation on $[n]$.
We remark that there is a natural extension of this scheme for mixnets with a
cascade of servers, each possessing its own decryption key \cite{cm06}. Indeed,
Camenisch and Mityagin use the Neff proof in order to verify the output of each
mix in the cascade. If a dishonest mix is detected, the protocol is run without
without using it.

\subsection{Decisional Diffie-Hellman} \label{sec-ddh}

Let $(p, q, g)$ be the public parameters of the ElGamal cryptosystem. The game
the adversary $\advD$ plays is defined as follows: let $a, b, r \getsr
[q-1]$ and let $A = g^a$, $B = g^b$, $Y_0 = g^r$, and $Y_1 = g^{ab}$. Choose
a random bit $b \getsr \bits$. On input $(A, B, Y_b)$, the adversary outputs a
bit $b^\prime \in \bits$ and wins if $b=b^\prime$. The advantage of $\advD$ is
defined as
\[
  \Delta(q) = |\Pr[\advD(A, B, Y_1)=1] - \Pr[\advD(A, B, Y_0)=1]|
\]
The DDH assumption is that, for every polynomial-time adversary $\advD$, the
function $\Delta(q)$ is negligible.

\subsection{Honest verifier zero-knowledge}

Zero-knowledge is defined in \cite{goldreich} with respect to the class of
formal languages with interactive proof systems. A language $L$ has an
interactive proof system if there exists a pair of probabilistic algorithms $(P,
V)$ such that for all $x \in L$, $\Pr[\brackets{P, V}(x) = 1] \ge 2/3$
(this property is called \emph{completeness}), and for all
$x \not\in L$, $\Pr[\brackets{P, V}(x) = 1] \le 1/3$ (this property is called
\emph{soundenss}). The class of languages
with interactive proof systems is denoted $\class{IP}$.

Let $L \in \class{IP}$
exhibited by $(P, V)$. We define security with respect to a game associated to
adversary $D$, verifier $V^*$, and simulator $M^*$. Suppose that $L$ is a finite
set\footnote{This is without loss of generality, since we
consider polynomial-time adversaries.} and let $x \getsr L$. Let $y_1 \getsr
\brackets{P, V^*}(x)$ and $y_0 \getsr M^*(x)$ choose a bit $b \getsr \bits$.
On input $(x, y_b)$, adversary $\advD$ outputs a bit $b^\prime$ and wins if
$b=b^\prime$. Let $k = |L|$ and define the advantage of $\advD$ as
\[
  \Delta(k) = |\Pr[\advD(x, y_1)=1] - \Pr[\advD(x, y_0)=1]|
\]
We say that $(P, V)$ is \emph{computational zero-knowledge} (or just
\emph{zero-knowledge}) if
for every probabilistic polynomial-time adversary $\advD$ and
  for every probabilistic polynomial-time verifier $V^*$,
    there exists probabilistic polynomial-time simulator $M^*$ such that $\Delta(k)$ is
    negligible.
The security notion achieved by Neff's protocol is somewhat weaker: a proof system $(P, V)$ is
\emph{honest verifier zero-knowledge} (HVZK) if
for every probabilistic polynomial-time adversary $\advD$,
  there exists a probabilistic polynomial-time simulator $M^*$ such that
\[
  |\Pr[\advD(x, \brackets{P,V}(x))=1] - \Pr[\advD(x, M^*(x))=1]|
\]
is a negligible function of $k$.

\section{The Neff shuffle} \label{section:neff-shuffle}

We now give the syntax of an interactive proof system for mixnets. The prover
$P$ (nominally the mix server) possesses the secret key $\sk$ and the verifier
$V$ possesses the public key $\pk$. The common input of $P$ and $V$ is all of
the messages passed through the mix server. The mix takes as input a sequence of
ciphertexts $(\varc_1, \varr_1), \ldots, (\varc_n, \varr_n) \in \Zmodp^* \by
\Zmodp^*$ and outputs a sequence of plaintexts $(\varm_{1}, S_1), \ldots,
(\varm_{n}, S_n) \in
\Zmodp^* \by \Zmodp^*$. Let $x \in \bits^*$ denote the encoding of the input ciphertexts and
output plaintexts. We say $x \in L$ if and only if there exists a bijection $\pi
: [n] \rightarrow [n] $ such that for every $i \in [n]$,
$(\varm_{\pi(i)}, S_{\pi(i)}) = \dec_k(\varc_{i}, \varr_{i})$.

A trivial proof system for $L$ is one in which the prover transmits to the
verifier its secret key and the verifier decrypts each of the ciphertexts. Of
course, this completely exposes $\pi$. Our goal is to specify a proof system
$(P, V)$ in which $P$ convinces $V$ of the existence of $\pi$ without revealing
the permutation or the secret key. A proof system that is HVZK is sufficient to
achieve this goal, which we present in section~\ref{sec-shuffle}. This protocol
is built up from simpler protocols presented in sections~\ref{sec-ilmp}
and~\ref{sec-shuffle0}. Finally, in section~\ref{sec-mix-shuffle}, we show how
to apply the main protocol to mixnets.

\subsection{The basic protocol and its security} \label{sec-ilmp}
The main building block for the protocol is a zero-knowledge proof for the
\emph{iterated logarithmic multiplication problem} (ILMP).
Let $n \in \N$, $x_1, \ldots, x_n \in [q-1]$, and $y_1, \ldots y_n \in [q-1]$.
Let $X_i = g^{x_i}$ and $Y_i = g^{y_i}$ for each $i \in [n]$. Given the
sequences $(X_1, \dots X_n)$ and $(Y_1, \dots, Y_n)$, the problem is to decide
if $g^{x_1 \cdots x_n} = g^{y_1 \cdots y_n}$.
Neff points out this problem is closely related to the Chaum-Pederson signature
scheme \cite{cp92} and the corresponding solution is nothing more than a
multidimensional generalization of their proof.
The following is an interactive proof for ILMP.
\begin{figure}[h]
  \oneCol{0.80}{
  \underline{$\textsc{Silmpp}_{\brackets{P,V}}\big( (X_1, \cdots, X_n), (Y_1, \dots,  Y_n)\big)$}:
  \begin{enumerate}
    \vspace{4pt}
    \item[P1.]
      Let $\theta_0 = \theta_1 = 0$ and $\theta_1, \ldots, \theta_{n-1} \getsr [q-1]$.
      Let $A_i = X_i^{\theta_{i-1}} \cdot Y_i^{\theta_i}$ for every $i \in [n]$.
      Send the sequence $A_1, \ldots A_n$ to $V$.

    \vspace{4pt}
    \item[V1.]
      Let $\gamma \getsr [q-1]$. Send $\gamma$ to $P$.

    \vspace{4pt}
    \item[P2.]
      Compute $r_1, \ldots, r_{n-1} \in [q-1]$ satisfying the set of equations:
      \begin{equation} \label{eq-ilmp}
        \begin{aligned}
           Y_1^{r_1} &= A_1 \cdot X_1^{-\gamma} \\
           X_i^{r_{i-1}} \cdot Y_i^{r_i} &= A_i \qquad (\forall \, i \in [2..n-1]) \\
           X_n^{r_{n-1}} &= A_n \cdot Y_n^{(-1)^{(n-1)}\gamma} \\
       \end{aligned}
     \end{equation}
     Send the sequence $r_1, \ldots r_{n-1}$ to $V$.

    \vspace{4pt}
    \item[V2.]
      Accept if equation~(\ref{eq-ilmp}) holds; otherwise reject.

  \end{enumerate}
  }
  \caption{SILMP protocol, an interactive proof for ILMP. The prover $P$ knows
    $x_1, \ldots x_n$ and $y_1, \ldots, y_n$.}
\end{figure}

\heading{Completeness.}
Assume that for every $i \in [n]$, it holds that $x_i \ne 0$ and $y_i \ne 0$.
(Otherwise, the verifier could see by inspection whether or not $(X_1, \ldots
X_n),(Y_1, \ldots, Y_n)$ is in the language. If $(\exists \, i \in
[n]) \, X_i = 1$, then $x_i = 0$ and $g^{x_1 \cdots x_n} = 1$. The same is true for
$Y_i$.)
Taking the log base-$g$ of both sides of equation~(\ref{eq-ilmp}) yields the
following system of linear equations:
\begin{equation}\label{eq-ilmp-completeness}
  \begin{bmatrix}
    y_1 & 0   & 0      & 0       & \cdots  & 0       \\
    x_2 & y_2 & 0      & 0       & \cdots  & 0       \\
    0   & x_3 & y_3    & 0       & \cdots  & 0       \\
        &     &        &         & \ddots  &         \\
    0   & 0   & \cdots & 0       & x_{n-1} & y_{n-1} \\
    0   & 0   & \cdots & 0       & 0       & y_{n}
  \end{bmatrix}
  \begin{bmatrix}
    r_1 - \theta_1 \\
    r_2 - \theta_2 \\
    r_3 - \theta_3 \\
    \vdots \\
    r_{n-2} - \theta_{n-2} \\
    r_{n-1} - \theta_{n-1}
  \end{bmatrix} =
  \begin{bmatrix}
    -\gamma x_1 \\
    0 \\
    0 \\
    \vdots \\
    0 \\
    (-1)^{(n-1)}\gamma
  \end{bmatrix}
\end{equation}
It can be shown that the $(n \times n)$ matrix in the left-hand side of
equation~(\ref{eq-ilmp-completeness}) is non-singular. This implies there exists
a unique solution $(r_1, \ldots, r_{n-1})$ satisfying this system. Neff gives
the solution explicitly:
\begin{equation} \label{eq-ilmp-findr}
  r_i - \theta_i = (-1)^{(n-i-1)} \gamma \prod_{j=i+1}^n {y_i}{x_i}^{-1}
\end{equation}
(The inverse of $x_i$ is computed in the group $\Zmodq^*$. This is
found using Euclid's algorithm.) Thus, if the prover isn't cheating, then the
verifier accepts with probability 1.

\heading{Soundness.}
Suppose the prover is cheating and $g^{x_1 \cdots x_n} \ne g^{y_1 \cdots y_n}$.
For any $r_1, \ldots, r_{n-1} \in [q-1]$, there is at most one $\gamma \in
[q-1]$ such that equation~(\ref{eq-ilmp}) holds. If $\gamma$ is chosen randomly,
then the verifier accepts with probability at most $1/q$.

\heading{Security.}
Let $x$ be an instance of ILMP.
The claim is that if the DDH assumption holds, then for every probabilistic
polynomial $\advD$, there exists a probabilistic, polynomial time simulator
$M^*$ such that
\[
  \Delta(q) = |\Pr[D(x, \brackets{P,V}(x))=1] - \Pr[D(x, M^*(x))=1|
\]
is negligible. To prove this claim, one would argue the contrapositive: suppose
there exists adversary $\advD$ such that for every simulator $M^*$, its
advantage in distinguishing the distribution $\brackets{P,V}(x)$ and $M^*(x)$ is
non-negligible. Then there exists an adversary $\advA$ whose advantage in the DDH
game outlined in section~\ref{sec-ddh} is non-negligible. We omit the details of
the proof here, but claim this can be shown via a reduction from $\advA$ to
$\advD$.

\subsection{Simple $n$-shuffle} \label{sec-shuffle0}
Let $n \in \N$, $x_1, \ldots, x_n \in [q-1]$, and $y_1, \ldots y_n \in [q-1]$.
Let $X_i = g^{x_i}$ and $Y_i = g^{y_i}$ for each $i \in [n]$. Let $c, d \in
[q-1]$, $C = g^c$, and $D = g^d$. We say that $x \in \textsc{Shuffle0}$ if $x$
is an encoding of $(X_1, \dots X_n)$, $(Y_1, \dots Y_n)$, $C$, and $D$ and there
exists a permutation $\pi : [n] \to [n]$ such that for every $i \in [n]$, it
holds that $Y_i^d = X_{\pi(i)}^c$. A zero-knowledge proof for
$\textsc{Shuffle0}$ is specified in figure~\ref{fig-shuffle0}.
\begin{figure}[h]
  \oneCol{0.80}{
  \underline{$\textsc{Shuffle0}_{\brackets{P,V}}\big( (X_1, \cdots, X_n), (Y_1, \dots,  Y_n), C, D\big)$}:
  \begin{enumerate}
    \vspace{4pt}
    \item[V1.] Let $t \getsr [q-1]$, $U = D^t$, and $W = C^t$. Let $\hat{X}_i =
      X_i \cdot U^{-1}$ and $\hat{Y}_i = Y_i \cdot W^{-1}$ for every $i \in [n]$. Send $t$ to $P$.

    \vspace{4pt}
    \item[P1.] Let $U = g^{dt}$ and $W = g^{ct}$. Let $\hat{X}_i = X_i \cdot U^{-1}$ and
      $\hat{Y}_i = Y_i \cdot W^{-1}$ for every $i \in [n]$. Run
      \[\textsc{Silmpp}_{\brackets{P,V}}\big( (\hat{X}_1, \dots, \hat{X}_n, C,
      \dots, C), (\hat{Y}_1, \dots, \hat{Y}_n, D, \dots, D) \big)\]
      where the number of $C$'s and $D$'s is $n$.

    \vspace{4pt}
    \item[V2.] Accept if $V$ accepts in step P1.
  \end{enumerate}
  }
  \caption{The simple $n$-shuffle, an interactive proof system for
    $\textsf{Shuffle0}$. The prover knows $x_1, \ldots x_n$ and $y_1, \ldots,
    y_n$, $c$, and $d$. (Here, inverse is computed in $\Zmodp^*$.)
  }
  \label{fig-shuffle0}
\end{figure}

\subsection{General $n$-shuffle} \label{sec-shuffle}
The simple $n$-shuffle is \emph{almost} what we need; its only limitation is
that it is required that the prover knows the log of each $X_i$ and $Y_i$. The
general $n$-shuffle, outlined in this section, overcomes this deficiency.
Let $X_1, \dots X_n, Y_1, \dots, Y_n \in \Zmodp^*$, $k \in [q-1]$, and $K =
g^k$. We say $(Y_1, \dots, Y_n)$ is a shuffling of $(X_1, \dots X_n)$ if there
exists a permutation $\pi : [n] \to [n]$ such that for every $i \in [n]$, it
holds that $Y_i = X_{\pi(i)}^k$. We say $x \in \textsc{Shuffle}$ if $x$ is an
an encoding of $X_1, \dots X_n, Y_1, \dots, Y_n$, and $K$ where $Y_1, \dots Y_n$ is a
shuffling of $X_1, \dots, X_n$. The protocol in
figure~\ref{fig-shuffle} specifies a zero-knowledge proof for \textsc{Shuffle}.
\begin{figure}[h]
  \oneCol{0.80}{
  \underline{$\textsc{Shuffle}_{\brackets{P,V}}\big( (X_1, \cdots, X_n), (Y_1,\dots,  Y_n), K\big)$}:
  \begin{enumerate}
    \vspace{4pt}
    \item[P1.] Let $e_{1,1}, \dots, e_{1,n}, e_{2,1}, \dots, e_{2,n}, d \getsr
      [q-1]$. For each $i \in [n]$, let $E_{1,i} = g^{e_{1,i}}$, $E_{2,i} =
      g^{e_{2,i}}$, and $D = g^d$. Send $(E_{1,1}, \dots E_{1,n})$, $(E_{2,1}, \dots
      E_{2,n})$, and $D$ to $V$.

    \vspace{4pt}
    \item[V1.] Let $f_{1,1}, \dots, f_{1,n}, f_{2,1}, \dots, f_{2,n} \getsr
      [q-1]$ and send $(f_{1,1}, \dots f_{1,n})$ and $(f_{2,1}, \dots
      f_{2,n})$ to $P$.

    \vspace{4pt}
    \item[P2.] For every $j \in [2]$ and $i \in [n]$, let $\alpha_{j,i} = d \cdot
      f_{j,\pi^{-1}(i)} \cdot e_{j,\pi^{-1}(i)}$,  and $F_{j,i} = g^{\alpha_{j,i}}$.
      Send $F_{1,1}, \dots,  F_{1,n}$ and $F_{2,1}, \dots, F_{2,n}$ to $V$.

    \vspace{4pt}
    \item[V2.] Let $\gamma \getsr [q-1]$. Send $\gamma$ to $P$.

    \vspace{4pt}
    \item[P3.] For every $j \in [2]$ and $i \in [n]$,
      let $\beta_{j,i} = f_{j,i} \cdot e_{j,i}$,
      $U_i = g^{\beta_{1,i} + \gamma \beta_{2,i}}$ and
      $V_i = g^{\alpha_{1,i} + \gamma \alpha_{2,i}}$.
      Run
      \[
        \textsc{Shuffle0}_{\brackets{P,V}}\big( (U_1, \dots, U_n), (V_1, \dots,
        V_n), g, D \big)
      \]

    \vspace{4pt}
    \item[P4.]
      Let $a_1, \dots, a_n, b_1, \dots, b_{n-1} \getsr [q-1]$ and
      \[
        b_n = v_n^{-1} \Big( \sum_{i=1}^{n-1} b_i v_i - d \sum_{i=1}^{n-1} a_i
        u_i \Big)
      \]
      where $u_i = \log U_i$ and $v_i = \log V_i$. For every $i \in [n]$, let
      $A_i = g^{a_i}$ and $B_i = g^{b_i}$. Send $(A_1, \dots, A_n)$ and $(B_1,
      \dots, B_n)$ to $V$.

    \vspace{4pt}
    \item[V3.] Let $\lambda \getsr [q-1]$. Send $\lambda$ to $P$.

    \vspace{4pt}
    \item[P5.] For every $i \in [n]$, let $s_i = a_i + \lambda u_i$,
      $r_i = b_i + \lambda v_i$, $P = \prod_{i=1}^n X_i^{r_i}$, and
      $Q = \prod_{i=1}^n Y_i^{s_i}$. Send $P$ and $Q$ to $V$.

    \vspace{4pt}
    \item[P6.] Run $\textsc{Silmpp}_{\brackets{P,V}}\big( (P,Q), (D,K) \big)$

    \vspace{4pt}
    \item[V4.] Accept if $V$ accepts in steps P3 and P6 and for every $i \in N$,
      $g^{s_i} = A_i \cdot U_i^\lambda$ and
      $g^{r_i} = B_i \cdot V_i^\lambda$.

  \end{enumerate}
  }
  \caption{The general $n$-shuffle, an interactive proof system for
    $\textsf{Shuffle}$. The prover knows the secret key  $k$ and the permuation
    $\pi$.
  }
  \label{fig-shuffle}
\end{figure}

\subsection{Applying the protocol to mixnets} \label{sec-mix-shuffle}
Consider the following algorithm for a mixnet constructed from the ElGamal
encryption scheme:
\newcommand{\ind}{\hspace{12pt}}
\begin{tabbing}
  \underline{Algorithm $\textsc{Mix}_k\big( (C_i, R_i)_{i=1}^n \big)$}\\[4pt]
  \ind $\pi \getsr \Perm(n)$\\
  \ind For $i \gets 1$ to $n$ do \\
  \ind\ind $(M_i, S_i) \gets \dec_k(C_{\pi(i)}, R_{\pi(i)})$ \\
  \ind Return $(M_i, S_i)_{i=1}^n$
\end{tabbing}
Let $K = g^k$ denote the corresponding public key.
For every $i \in [n]$, we have that $S_i = R_{\pi(i)}^k$ and $M_i = C_{\pi(i)}
\cdot S_i^{-1}$. To prove that the mix didn't cheat, that is $C_{i}$ is the ciphertext
corresponding to the message $M_{\pi(i)}$, it suffices to prove $S_i =
R_{\pi(i)}^k$. This is nothing more than an instance of $\textsc{Shuffle}$. The
output of the mix can be verified by running
$\textsc{Shuffle}_{\brackets{P,V}}\big( (R_1, \dots, R_n), (S_1, \dots, S_n), K
\big)$.

\section{Implementation} \label{sec-implementation}
I implemented the simple $n$-shuffle of section~\ref{sec-shuffle0} in Go. The
code is available online at \textit{github.com/cjpatton/shuffle}. I wrote
a pair of functions \texttt{Shuffle0Prove} and \texttt{Shuffle0Verify}, which
implement the prover and verifier roles of the $\textsc{Shuffle0}$ proof
respectively. They use a standard Go channel for communication. This makes it
convenient to adapt the code for a real-world application using a network socket
for example.

My goal for this project was to implement the general $n$-shuffle and adapt it
for the mixnet setting; I skipped it because of time constraints. Coding this
algorithm ended up being technically difficult. In particular, there were a
number of details about the computations left implicit in the paper that I had
to figure out. For example, the system of equations in equation~(\ref{eq-ilmp})
in my implementation differ slightly from those presented in the paper, so I
ended up doing some math-hacking to get it to work. Once I got \textsc{Silmpp}
straightened out, implementing \textsc{Shuffle0} was pretty straight-forward.

\begin{thebibliography}{99}

  \bibitem{chaum81} David Chaum, 1981. ``Untraceable electronic mail, return addresses, and digital pseudonyms.''
    {\tt http://freehaven.net/anonbib/cache/chaum-mix.pdf}

  \bibitem{elgamal84} Taher Elgamal, ``A public key cryptosystem and a signature
    scheme based on discrete logarithms.'' Appeared in \textit{CRYPTO'84.}

  \bibitem{cp92}  D. Chaum and T.P. Pedersen. ``Wallet databases with
    observers.'' Appeared in \textit{CRYPTO'92.}

  \bibitem{ty98} Yiannis Tsiounis and Moti Yung 1998. ``On the security of
    ElGamal based encryption.'' Appeared in \textit{PKC'98.}

  \bibitem{goldreich} Oded Goldreich, 2001. \textit{Foundations of Cryptography.}

  \bibitem{neff01}
    Andrew Neff. "A verifiable secret shuffle and its application to e-voting."
    Proceedings of the 8th ACM conference on Computer and Communications Security.
    ACM, 2001.

  \bibitem{cm06} Jan Camenisch and Anton Mityagin, 2006. ``Mix-network with
    stronger security.''

\end{thebibliography}

\end{document}
